---
title: "Generalism data results"
author: "Josephine Walker"
date: "14 May 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo=FALSE,warning=FALSE,error=FALSE,message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Prediction 1: It is easier for generalist endoparasites to invade if the available hosts are large-bodied
I looked for a relationship between the mean, coefficient of variation ($\sigma/\mu$) and maximum MaxL (fish host length) value for each parasite’s hosts and each of the four parasite generalism indices (*degree*, the number of hosts; *gen*, a binary of whether the parasite has more than one host; *S~TD~*, the average taxonomic distance between the hosts; and *VarS~TD~*, the variance of *S~TD~*).

```{r maxl,echo=FALSE,warning=FALSE}
load("~/dat/generalismtheory/hp.full.simple.RData")

hp.endo <- hp.full.simple %>% filter(Endoparasite =="Endo")
hp.endo$gen <- FALSE
hp.endo$gen[hp.endo$degree > 1] <- TRUE



# endolong <- hp.endo %>% gather("index","value",10:13)
# ggplot(data=endolong,aes(x=meanMaxL,y=value)) + geom_jitter() + theme_bw() + facet_wrap(~index,nrow=2,scales="free") + ylab("index value")
# mnMaxLdeg <- with(hp.endo,cor.test(meanMaxL,degree,method="spearman"))
# mnMaxLgen <- with(hp.endo,cor.test(meanMaxL,as.numeric(gen),method="spearman"))
# mnMaxLstd <- with(hp.endo,cor.test(meanMaxL,S_TD,method="spearman"))
# mnMaxLvstd <- with(hp.endo,cor.test(meanMaxL,VarS_TD,method="spearman"))
# hp.endo %>% group_by(gen) %>% summarize(mean=mean(meanMaxL,na.rm=T),median=median(meanMaxL,na.rm=T))

hp.endoDirect <- hp.endo %>% filter(Complex=="No")

endoDlong <- hp.endoDirect %>% gather("index","value",10:13)
# ggplot(data=endoDlong,aes(x=meanMaxL,y=value)) + geom_jitter() + theme_bw() + facet_wrap(~index,nrow=2,scales="free") + ylab("index value") + geom_smooth(method="lm")
# ggplot(data=endoDlong,aes(x=maxMaxL,y=value)) + geom_jitter() + theme_bw() + facet_wrap(~index,nrow=2,scales="free") + ylab("index value") + geom_smooth(method="lm")

endoDlong2 <- endoDlong %>% gather("length_measure","length",7:9)
ggplot(data=endoDlong2,aes(x=length,y=value)) + geom_jitter() + theme_bw() + facet_grid(length_measure~index,scales="free") + ylab("index value") + geom_smooth(method="lm")

endoDlong$cvmaxL <- endoDlong$sdMaxL/endoDlong$meanMaxL
ggplot(data=endoDlong,aes(x=cvmaxL,y=value)) + geom_jitter() + theme_bw() + facet_wrap(~index,nrow=2,scales="free") + ylab("index value") + geom_smooth(method="lm")

# summary(lm(hp.endoDirect$VarS_TD ~ hp.endoDirect$cvmaxL))
# summary(lm(hp.endoDirect$S_TD ~ hp.endoDirect$sdMaxL))
# cor.test(hp.endoDirect$S_TD, hp.endoDirect$sdMaxL,method="spearman")
# cor.test(hp.endoDirect$VarS_TD, hp.endoDirect$maxMaxL,method="spearman")
```

## Ectoparasites and host body size - more complex relationship expected

```{r ecto, echo=FALSE}
hp.ecto <- hp.full.simple %>% filter(Endoparasite =="Ecto")
hp.ecto$gen <- FALSE
hp.ecto$gen[hp.ecto$degree > 1] <- TRUE
hp.ectoDirect <- hp.ecto %>% filter(Complex=="No")

# divide into quartiles
hp.ectoDirect$mean.lcat <- 1
hp.ectoDirect$mean.lcat[hp.ectoDirect$meanMaxL > 27] <- 2
hp.ectoDirect$mean.lcat[hp.ectoDirect$meanMaxL > 54] <- 3
hp.ectoDirect$mean.lcat[hp.ectoDirect$meanMaxL > 124.5] <- 4

ectoDlong <- hp.ectoDirect %>% gather("index","value",10:13)
ectoDlong2 <- ectoDlong %>% gather("length_measure","length",7:10)

ectoDlong2 %>% filter(length_measure=="meanMaxL") %>% ggplot(data=.,aes(x=length,y=value)) + geom_jitter() + theme_bw() + facet_wrap(~index,scales="free") #+ ylab("index value") #+ geom_smooth(method="lm") #+ scale_y_log10() + scale_x_log10()

ectoDlong2 %>% filter(length_measure=="mean.lcat") %>% ggplot(data=.,aes(x=as.factor(length),y=value)) + geom_boxplot() + theme_bw() + facet_wrap(~index,scales="free") + ylab("index value")

ectoDlong$cvmaxL <- ectoDlong$sdMaxL/ectoDlong$meanMaxL
ggplot(data=ectoDlong,aes(x=cvmaxL,y=value)) + geom_jitter() + theme_bw() + facet_wrap(~index,nrow=2,scales="free") + ylab("index value") + geom_smooth(method="lm")

ectoDlong2 %>% filter(length_measure=="mean.lcat") %>% group_by(index, length) %>% summarize(mean=mean(value,na.rm=T),median=median(value,na.rm=T))

with(hp.ectoDirect,kruskal.test(degree ~ as.factor(mean.lcat)))
lm1 <- lm(data=hp.ectoDirect,degree ~ as.factor(mean.lcat))
summary(lm1)
plot(lm1)
lm2 <- glm(data=hp.ectoDirect,degree ~ as.factor(mean.lcat),family="poisson") # or quasipoisson
summary(lm2)
plot(lm2)
lm3 <- lm(data=hp.ectoDirect,S_TD ~ as.factor(mean.lcat))
summary(lm3)
plot(lm3)

ggplot(data=hp.ectoDirect,aes(degree)) + geom_histogram(binwidth=2) + theme_bw() + facet_wrap(~mean.lcat)

library(MASS)
m1 <- glm.nb(data=hp.ectoDirect,degree ~ as.factor(mean.lcat))
m0 <- glm.nb(data=hp.ectoDirect,degree ~ 1)

# library(VGAM)
# m2 <- vglm(data=hp.ectoDirect,degree ~ mean.lcat,family=posnegbinomial()) # errorrr

```
Negative binomial GLM seems like the best way to compare Ecto degree to quartiles of length categories. The estimates are log values, so exp() to get the estimate. Except, it's strictly positive so should use a zero truncated model?


Index isn't calculated based on geography, is it?
```{r geo, echo=FALSE}
load("~/dat/generalismtheory/hp.full.geo.RData")
hp.full.geo$gen <- FALSE
hp.full.geo$gen[hp.full.geo$degree > 1] <- TRUE
hp.geoDirect <- hp.full.geo %>% filter(Complex=="No")
geoDlong <- hp.geoDirect %>% gather("index","value",11:14)

geoDlong %>% ggplot(data=.,aes(x=GEO,y=value)) + geom_violin() + theme_bw() + facet_wrap(~index,scales="free") #+ ylab("index value") #+ geom_smooth(method="lm") #+ scale_y_log10() + scale_x_log10()
```

```{r site,echo=FALSE}
hp.site <- hp.full.simple %>% gather("index","value",10:12) %>% filter(Complex=="No")
  
hp.site %>% ggplot(data=.,aes(x=Endoparasite,y=value)) + geom_violin() + theme_bw() + facet_wrap(~index,scales="free") 

```

```{r complex,echo=FALSE}
hp.comp <- hp.full.simple %>% gather("index","value",10:12) %>% filter(Complex == "Yes" | Complex=="No")
  
hp.comp %>% ggplot(data=.,aes(x=Complex,y=value)) + geom_boxplot() + theme_bw() + facet_wrap(~index,scales="free") 

```

```{r trophic,echo=FALSE}
hp.trop <- hp.full.simple %>% gather("index","value",10:12) 
  
hp.trop %>% ggplot(data=.,aes(x=Trophic,y=value)) + geom_boxplot() + theme_bw() + facet_wrap(~index,scales="free") 

```

Maybe all the missing data is biased because it's for parasite species that are less well studied...