---
title: "Generalism Index"
author: "Josephine Walker"
date: "March 27, 2016"
output: pdf_document
---

The aim of this document is to calculate measures of generalism from the fish database, and assess the distribution of generalism scores by parasite characteristics.

The [preliminary] data are from the Fish Parasite Ecology Database with parasite characteristics added by Amy Ellison and Jo Cable (http://www.esapubs.org/archive/ecol/E094/045/metadata.php)

```{r,cache=TRUE,results="hide"}
# read in data & remove duplicates

library(dplyr,suppressPackageStartupMessages())

setwd("C:/Users/Josephine/Google Drive/Conferences and Courses/Transmission retreat")

dat <- read.csv("fishparasites.csv")

u <- dat %>% select(P_Genus,P_species,H_Genus,H_species) %>% duplicated(MARGIN=1) %>% which()
dat.u <- dat[-u,]


# to clean data - should check list of parasite and host species as some spellings are very similar, are these really different species?
# also values in each characteristic column are not consistent, ie yes/Yes
names(dat.u)
levels(dat.u$Host_stage)
levels(dat.u$Host_stage)[5] <- "Intermediate"
levels(dat.u$Endoparasite)
levels(dat.u$Endoparasite)[4:6] <- "Endoparasite"
levels(dat.u$Gender)
levels(dat.u$Gender) <- c("","Dioecious","Dioecious","Hermaphrodite","Hermaphrodite","Hermaphrodite")
levels(dat.u$Complex.)
levels(dat.u$Complex.) <- c("","",NA,"Simple","Complex","Complex","Complex","Complex","Both")
levels(dat.u$Sexual.)
levels(dat.u$Sexual.) <- c("",NA,"Both","Both","Oviparous","Sexual","Sexual","Sexual")
levels(dat.u$Reproduction)
levels(dat.u$Reproduction) <- c("",NA,"Oviparous","Oviparous","Oviparous","Ovoviviparous","Ovoviviparous","Sexual","Viviparous") # sexual shouldn't be here?
levels(dat.u$Host_no.) # how should this be categorized?
levels(dat.u$Environment_LHS.)
levels(dat.u$Environment_LHS.) <- c("","No","Yes","Yes")
levels(dat.u$Env_mode.)
levels(dat.u$Env_mode.)[4:6] <- "Water" # how should this be categorized?
levels(dat.u$Motile_LHS.)
levels(dat.u$Motile_LHS.) <- c("","No","No","Yes","Yes")
levels(dat.u$Vector.) 
levels(dat.u$Vector.) <- c("","No","No","Yes")
levels(dat.u$Trophic.)
levels(dat.u$Trophic.) <- c("",NA,NA,"No","No","Yes","Yes")
levels(dat.u$Horizontal.) # Are these all "yes"?

## host taxa
levels(dat.u$H_O)
levels(dat.u$H_O)[15:17] <- "Characiformes"
levels(dat.u$H_O)[c(18,20,21)] <- "Cyprinodontiformes"
levels(dat.u$H_O)[26:27] <- "Gonorhynchiformes" 
levels(dat.u$H_O)[33:34] <- "Lepidosireniformes"
levels(dat.u$H_O)[46:47] <- "Perciformes"
# there are many more to be corrected at lower levels - indices will change if species are added so I won't do this now

```

Generalism of the parasites can be calculated in a variety of ways, and may take into account relative prevalence, abundance and/or relatedness between host species [Poulin & Mouillot 2003, 2005]. Other indices account for connectedness within a host-parasite network and may account for prevalence or abundance [Dormann].  

In this data set, we don't have information on number of parasites per host or on prevalence or study effort, only reported associations between hosts and parasites. As a result we calculate only:

* **Degree**: the number of host species a parasite is associated with
* **$S_{TD}$**: the average taxonomic distinctness of all host species used by a parasite species [Poulin & Mouillot 2003]: $$S_{TD}=2[s(s-1)]^{-1}\sum \sum_{i<j} \omega_{ij}$$

    $S_{TD}$ can only be calculated for parasites infecting more than one host species, those infecting one species are assigned a value of 1. In addition $VarS_{TD}$ can only be calculated for hosts infecting more than 2 host species:
$$VarS_{TD}=[s(s-1)]^{-1}\sum \sum_{i\neq j}\left (\omega_{ij}-S_{TD}  \right )^2$$
    But this is the same as: $$VarS_{TD}=2[s(s-1)]^{-1}\sum \sum_{i< j}\left (\omega_{ij}-S_{TD}  \right )^2$$





* Another method could use host phylogenetic distance (eg from time tree of life) instead of categorized phylogenies
    + Do we have data for all species?

```{r,cache=TRUE}
library(ape,suppressPackageStartupMessages())

tree <- read.tree("C:/Users/Josephine/Google Drive/Thesis/Chapter 3 (Community Ecology)/ch3 stuff/TimetreeOfLife2015.nwk")

H_bin <- paste(dat.u$H_Genus,dat.u$H_species,sep="_") %>% unique()

pruned.tree<-drop.tip(tree, setdiff(tree$tip.label, H_bin))

# calculate path distance between each species (this should be time since last common ancestor*2)
pathlength <- cophenetic.phylo(pruned.tree)

paste("number of host species in database:",length(H_bin),sep=" ")
paste("number of nodes in pruned tree:", length(pruned.tree$tip.label),sep=" ")


```


Divergence time data are only available for about a quarter of the host species in the fish database. 

How many parasite species only infect one host in the database?

```{r,cache=TRUE}
dat.u$P_bin <- paste(dat.u$P_Genus,dat.u$P_species,sep="_")
dat.u$H_bin <- paste(dat.u$H_Genus,dat.u$H_species,sep="_")
degree <- dat.u %>% group_by(P_bin) %>% summarize(degree=n_distinct(H_bin),P_Taxon=first(P_Taxon),Host_stage=first(Host_stage),Endoparasite=first(Endoparasite),Gender=first(Gender),Complex=first(Complex.),Sexual=first(Sexual.),Reproduction=first(Reproduction),Host_no=first(Host_no.),Environment_LHS=first(Environment_LHS.),Env_mode=first(Env_mode.),Motile_LHS=first(Motile_LHS.),Vector=first(Vector.),Trophic=first(Trophic.),Horizontal=first(Horizontal.))
sum(degree$degree == 1)/nrow(degree)
quantile(degree$degree)

hist(log(degree$degree))
```

Does degree vary by taxonomic group or other parasite characteristics?

```{r,cache=TRUE}
boxplot(log(degree$degree) ~ degree$Gender)
boxplot(log(degree$degree) ~ degree$Host_stage)
boxplot(log(degree$degree) ~ degree$Endoparasite)
boxplot(log(degree$degree) ~ degree$Complex)
boxplot(log(degree$degree) ~ degree$Sexual)
boxplot(log(degree$degree) ~ degree$Reproduction)
boxplot(log(degree$degree) ~ degree$Host_no)
boxplot(log(degree$degree) ~ degree$Environment_LHS)
boxplot(log(degree$degree) ~ degree$Env_mode)
boxplot(log(degree$degree) ~ degree$Motile_LHS)
boxplot(log(degree$degree) ~ degree$Vector)
boxplot(log(degree$degree) ~ degree$Trophic)
boxplot(log(degree$degree) ~ degree$Horizontal)
boxplot(log(degree$degree) ~ degree$P_Taxon)
```

Calculate S_TD using taxonomic distance from data

```{r,eval=FALSE}

# calculate distance between each pair of host species (omega_ij) - original reference suggests up to phylum, we only have up to order. I was going to match orders to class but google revealed some controversy so will have to come back to this - any good sources for fish taxonomy?
dat.u$H_Class <- ""
dat.u$H_Phylum <- "Chordata"

hosts <- dat.u %>% select(H_O,H_F,H_Genus,H_species,H_Phylum,H_Class) %>% distinct()
hosts$H_bin <- paste(hosts$H_Genus,hosts$H_species,sep="_")
sum(table(hosts$H_bin) > 1) # some aren't unique presumably due to differences in capitalization etc at higher levels - check this again after data cleaning
# for now select unique based on H_bin
hosts.u <- hosts %>% distinct(H_bin)

# # http://stackoverflow.com/questions/17171148/non-redundant-version-of-expand-grid
expand.grid.unique <- function(x, y, include.equals=FALSE)
{
    x <- unique(x)

    y <- unique(y)

    g <- function(i)
    {
        z <- setdiff(y, x[seq_len(i-include.equals)])

        if(length(z)) cbind(x[i], z, deparse.level=0)
    }

    do.call(rbind, lapply(seq_along(x), g))
}

# write function to calculate S_TD
# only congeneric hosts will have value 1, so call one species only value 0?

get.S_TD <- function(dat) {
  S_TD <- 0
  VarS_TD <- NA
  
  if(nrow(dat) > 1) {
    temp <- data.frame(expand.grid.unique(dat$H_bin,dat$H_bin))
    temp$genus <- 0
    temp$family <- 0
    temp$order <- 0
    temp$class <- 0
    temp$phylum <- 0
    
    for (i in 1:nrow(temp)) {
      temp2 <- rbind(dat[dat$H_bin == temp$X1[i],],dat[dat$H_bin == temp$X2[i],])
      temp$genus[i] <- as.numeric(identical(temp2$H_Genus[1],temp2$H_Genus[2]))
      temp$family[i] <- as.numeric(identical(temp2$H_F[1],temp2$H_F[2]))
      temp$order[i] <- as.numeric(identical(temp2$H_O[1],temp2$H_O[2]))
      temp$class[i] <- as.numeric(identical(temp2$H_Class[1],temp2$H_Class[2]))
      temp$phylum[i] <- as.numeric(identical(temp2$H_Phylum[1],temp2$H_Phylum[2]))
    }

    temp$omega <- with(temp, 6 - genus - family - order - class - phylum)

    S_TD <- 2*(sum(temp$omega))/(nrow(dat)*(nrow(dat)-1))
    temp$Var <- (temp$omega - S_TD)^2 
    #print(temp)
    VarS_TD <- 2*(sum(temp$Var))/(nrow(dat)*(nrow(dat)-1))
  }
  return(data.frame(S_TD=S_TD,VarS_TD=VarS_TD,degree=nrow(dat)))
  #return(paste(S_TD,VarS_TD,sep="@"))
}

dat.u$H_O <- as.character(dat.u$H_O)
dat.u$H_F <- as.character(dat.u$H_F)
dat.u$H_Genus <- as.character(dat.u$H_Genus)

#test <- dat.u[1:100,] %>% group_by(P_bin) %>% do(get.S_TD(.))
# sub <- index[which(is.infinite(index$S_TD)),1]
# test2 <- dat.u[dat.u$P_bin %in% sub$P_bin,] %>% group_by(P_bin) %>% do(get.S_TD(.))


index <- dat.u %>% group_by(P_bin) %>% do(get.S_TD(.)) %>% ungroup()
save(index,file="index.RData")

````

The triangular scatter of the index matches what was observed in Poulin & Mouillot 2003.

```{r,cache=TRUE}
load("index.RData")
plot(x=index$degree,y=index$S_TD)
plot(x=log10(index$degree),y=index$S_TD)

#boxplots by parasite characteristic
fulldat <- full_join(degree,index,by=c("P_bin","degree"))
# dim correct

boxplot(fulldat$S_TD ~ fulldat$Gender)
boxplot(fulldat$S_TD ~ fulldat$Host_stage)
boxplot(fulldat$S_TD ~ fulldat$Endoparasite)
boxplot(fulldat$S_TD ~ fulldat$Complex)
boxplot(fulldat$S_TD ~ fulldat$Sexual)
boxplot(fulldat$S_TD ~ fulldat$Reproduction)
boxplot(fulldat$S_TD ~ fulldat$Host_no)
boxplot(fulldat$S_TD ~ fulldat$Environment_LHS)
boxplot(fulldat$S_TD ~ fulldat$Env_mode)
boxplot(fulldat$S_TD ~ fulldat$Motile_LHS)
boxplot(fulldat$S_TD ~ fulldat$Vector)
boxplot(fulldat$S_TD ~ fulldat$Trophic)
boxplot(fulldat$S_TD ~ fulldat$Horizontal)
boxplot(fulldat$S_TD ~ fulldat$P_Taxon)


```


