---
title: "Generalism data results"
author: "Josephine Walker"
date: "14 May 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo=FALSE,warning=FALSE,error=FALSE,message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Prediction 1: It is easier for generalist endoparasites to invade if the available hosts are large-bodied
I looked for a relationship between the mean, coefficient of variation ($\sigma/\mu$) and maximum MaxL (fish host length) value for each parasite’s hosts and each of the four parasite generalism indices (*degree*, the number of hosts; *gen*, a binary of whether the parasite has more than one host; *S~TD~*, the average taxonomic distance between the hosts; and *VarS~TD~*, the variance of *S~TD~*).

```{r maxl,echo=FALSE,warning=FALSE}
load("~/dat/generalismtheory/hp.full.simple.RData")

hp.endo <- hp.full.simple %>% filter(Endoparasite =="Endo")
hp.endo$gen <- FALSE
hp.endo$gen[hp.endo$degree > 1] <- TRUE
endolong <- hp.endo %>% gather("index","value",10:13)

ggplot(data=endolong,aes(x=meanMaxL,y=value)) + geom_jitter() + theme_bw() + facet_wrap(~index,nrow=2,scales="free") + ylab("index value")

mnMaxLdeg <- with(hp.endo,cor.test(meanMaxL,degree,method="spearman"))
mnMaxLgen <- with(hp.endo,cor.test(meanMaxL,as.numeric(gen),method="spearman"))
mnMaxLstd <- with(hp.endo,cor.test(meanMaxL,S_TD,method="spearman"))
mnMaxLvstd <- with(hp.endo,cor.test(meanMaxL,VarS_TD,method="spearman"))
# hp.endo %>% group_by(gen) %>% summarize(mean=mean(meanMaxL,na.rm=T),median=median(meanMaxL,na.rm=T))

hp.endoDirect <- hp.endo %>% filter(Complex=="No")

qplot(data=hp.endoDirect,x=meanMaxL,y=degree) + geom_smooth(method = "lm")
qplot(data=hp.endoDirect,x=maxMaxL,y=degree) + geom_smooth(method = "lm")


endoDlong <- hp.endoDirect %>% gather("index","value",10:13)
ggplot(data=endoDlong,aes(x=meanMaxL,y=value)) + geom_jitter() + theme_bw() + facet_wrap(~index,nrow=2,scales="free") + ylab("index value") + geom_smooth(method="lm")
ggplot(data=endoDlong,aes(x=maxMaxL,y=value)) + geom_jitter() + theme_bw() + facet_wrap(~index,nrow=2,scales="free") + ylab("index value") + geom_smooth(method="lm")

endoDlong2 <- endoDlong %>% gather("length_measure","length",7:9)
ggplot(data=endoDlong2,aes(x=length,y=value)) + geom_jitter() + theme_bw() + facet_grid(length_measure~index,scales="free") + ylab("index value") + geom_smooth(method="lm")

endoDlong$cvmaxL <- endoDlong$sdMaxL/endoDlong$meanMaxL
ggplot(data=endoDlong,aes(x=cvmaxL,y=value)) + geom_jitter() + theme_bw() + facet_wrap(~index,nrow=2,scales="free") + ylab("index value") + geom_smooth(method="lm")

summary(lm(hp.endoDirect$VarS_TD ~ hp.endoDirect$cvmaxL))
summary(lm(hp.endoDirect$S_TD ~ hp.endoDirect$sdMaxL))
cor.test(hp.endoDirect$S_TD, hp.endoDirect$sdMaxL,method="spearman")
cor.test(hp.endoDirect$VarS_TD, hp.endoDirect$maxMaxL,method="spearman")
```

## Large-bodied hosts are more likely to be infected by generalist than specialist endoparasites
and maximum and CV - sd/mean? `r mnMaxLdeg$estimate`

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)

load("~/dat/generalismtheory/hp.full.geo.stage.RData")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
