---
title: "Is generalism of fish macroparasites associated with parasite and/or host traits?"
author: "Josephine Walker"
date: "8 April 2016"
output: pdf_document
---

The aim of this document is to calculate measures of generalism from the fish database, and assess the distribution of generalism scores by parasite characteristics.

The data are from the Fish Parasite Ecology Database (http://www.esapubs.org/archive/ecol/E094/045/metadata.php) with crustacean parasites and parasite traits for all species added by Amy Ellison, Jo Cable and team

```{r,eval=FALSE}
# read in data & remove duplicates

library(dplyr,suppressPackageStartupMessages())
library(tidyr,suppressPackageStartupMessages())

# set working directory to data location
setwd("~/dat/generalismtheory")

dat <- read.csv("fishparas.csv",header=T)


# update levels for consistency
levels(dat$Host_stage) <- c("Both","Both","Definitive","Definitive","Intermediate","nd")
levels(dat$GEO) <- c(NA,"AFR","ANT","AUS","IND",NA,NA,"NEA","NEO","PAL",NA,NA)
# correct a few species names - haven't checked all only newly added ones
dat$H_species[dat$H_species=="insignisC"] <- "insignis"
dat$H_species[dat$H_species=="Coeruleus"] <- "coeruleus"
dat$H_species[dat$H_species=="crocodiluscrocodilus"] <- "crocodilus"
dat$H_species[dat$H_species=="carangis"] <- "carangus"

dat$pname <- paste(dat$P_Genus,dat$P_species,sep="_")
dat$hname <- paste(dat$hgenus,dat$H_species,sep="_")
dat$hp <- paste(dat$hname,dat$pname,sep="+")

dat <- dat[-which(dat$hname=="Colossoma_macroponum"),]

# extract list of host-parasite associations & traits
# association traits: Host_stage, GEO
pairs <- dat %>% select(hname,pname) %>% distinct()

# clean Host_stage
pairs.t1 <- dat %>% select(hp,hname,pname,Host_stage) %>% distinct()
pairs.dup <- pairs.t1[duplicated(pairs.t1[,1:2]) | duplicated(pairs.t1[,1:2], fromLast=TRUE),] %>% arrange(hname,pname) # all doubles 

cleanHS <- function(d) {
  if(any(d$Host_stage=="nd")) {
    new <- as.character(d$Host_stage[d$Host_stage!="nd"])
    return(new)
  }
  if(any(d$Host_stage=="Both")) {return("Both")}
  if(any(d$Host_stage=="Intermediate") & any(d$Host_stage=="Definitive")) {return("Both")}
}

pairs.new <- pairs.dup %>% group_by(hname,pname) %>% do(newHS=cleanHS(.))
pairs.new$newHS <- as.character(pairs.new$newHS)
# put back together
pairs.t2 <- left_join(pairs.t1,pairs.new,by=c("hname","pname"))
pairs.t2$Host_stage[!is.na(pairs.t2$newHS)] <- pairs.t2$newHS[!is.na(pairs.t2$newHS)]
pairs.t <- distinct(pairs.t2)
pairs.t$newHS <- NULL

# come back to GEO
pairs.g1 <- dat %>% select(hp,hname,pname,GEO) %>% distinct()
pairs.g1$val <- 1
pairs.g2 <- spread(pairs.g1,key=GEO,value=val,fill=0)
pairs.traits <- inner_join(pairs.t,pairs.g2,by=c("hp","hname","pname"))

save(pairs.traits,file="pairs.traits.RData")

############

# extract list of parasite traits
levels(dat$Endoparasite) <- c("Both","Ecto","Endo","Endo",NA)
levels(dat$Gender) <- c("Dioecious","Hermaphrodite","Hermaphrodite",NA)
levels(dat$Complex) <- c(NA,NA,"No","Yes","Yes_No")
levels(dat$Sexual) <- c(NA,"Both",NA,"Sexual","Sexual","Sexual")
levels(dat$Reproduction)[1:2] <- c(NA,NA)
levels(dat$Host_no)[c(1,8)] <- c(NA,NA)
levels(dat$Environment_LHS)[1:2] <- c(NA,NA)
levels(dat$Env_mode) <- c(NA,"napp",NA,"Water","Water")
levels(dat$Motile_LHS) <- c(NA,NA,"No","Yes","Yes")
levels(dat$Vector) <- c(NA,NA,"No","No","Yes")
levels(dat$Trophic) <- c(NA,NA,NA,NA,"No","Yes")
levels(dat$Horizontal) <- c(NA,"Yes",NA,"Yes")
levels(dat$P_Taxon) <- c("A","A","C","C","CR","M","M","N","N","T","T")

para <- dat %>% select(pname,P_Taxon,Endoparasite,Gender,Complex,Sexual,Reproduction,Host_no,Environment_LHS,Env_mode,Motile_LHS,Vector,Trophic,Horizontal) %>% distinct() # distinct rows of parasite traits
para.dup <- para[duplicated(para$pname) | duplicated(para$pname,fromLast = TRUE),] %>% arrange(pname) # find duplicated parasites
write.csv(para.dup,file="para.dup.out.csv")


# merge back with new row as whole row replacing that parasite name
para.dup.in <- read.csv("para.dup.in.csv",stringsAsFactors = FALSE)

para.t <- anti_join(para,para.dup,by="pname")
para.traits <- rbind(para.t,para.dup.in) %>% arrange(pname)

save(para.traits,file="para.traits.RData")

############
# host traits
host <- dat %>% select(hname,MaxL,K,Y,Ym,T,F,B,M,AOO,LAT,LON) %>% distinct() %>% arrange(hname)
host.dup <- host[duplicated(host$hname) | duplicated(host$hname,fromLast = TRUE),] # not all doubles
host2 <- anti_join(host,host.dup,by="hname")

# a number of duplicates are the same row filling in
dub <- host.dup[c(2,42),-1]
host.dup2 <- anti_join(host.dup,dub,by=c("MaxL", "K", "Y", "Ym", "T", "F", "B", "M", "AOO", "LAT", "LON")) %>% arrange(hname)

host3 <- rbind(host2,host.dup2)

host.dup3 <- host.dup2[duplicated(host.dup2$hname) | duplicated(host.dup2$hname,fromLast = TRUE),]
write.csv(host.dup3,file="host.dup.out.csv",row.names = FALSE)

host.dup.in <- read.csv("host.dup.in.csv",stringsAsFactors = FALSE)
host.t <- anti_join(host3,host.dup3,by="hname")
host.traits <- rbind(host.t,host.dup.in) %>% arrange(hname)
anyDuplicated(host.traits$hname)

save(host.traits,file="host.traits.RData")

#############
# host taxa - old list with class added plus new ones not on that list
# old class list
host.taxa1 <- read.csv("new_hostlist_wclass.csv",stringsAsFactors = FALSE)
#anyDuplicated(host.taxa$H_bin)
host.taxa2 <- host.taxa1 %>% select(-H_O.1,-uid)
host.taxa2$H_O[which(host.taxa2$H_O=="Tetradontiformes")] <- "Tetraodontiformes"
host.taxa2$H_F <- toupper(host.taxa2$H_F)
host.taxa2$H_O <- toupper(host.taxa2$H_O)
host.taxa3 <- distinct(host.taxa2)

# which hosts do we not have data on?
taxa <- dat %>% select(hname,H_O,H_F,hgenus,H_species) %>% distinct(hname) %>% arrange(hname)
taxa2 <- anti_join(taxa,host.taxa3,by=c("hname"="H_bin")) %>% arrange(hname)
#anyDuplicated(taxa2$hname)

write.csv(taxa2,file="newhostnames.csv")
new <- read.csv("newhostnames-class.csv") 
newtaxa <- new %>% select(-uid)
names(newtaxa) <- c("H_bin","H_O","H_F","H_Genus","H_species","H_class")
newtaxa$H_Phylum <- "Chordata"

# look at which ones from the old list aren't in new dat - mostly have capital & space issues that must have been corrected already
taxarev <- anti_join(host.taxa3,taxa,by=c("H_bin"="hname")) %>% arrange(H_bin) 

# exclude taxa which aren't in new dat
taxa3 <- anti_join(host.taxa3,taxarev,by=("H_bin"))

# group old and new
alltaxa <- bind_rows(newtaxa,taxa3)
names(alltaxa)[1] <- "hname"

taxa.dup <- alltaxa[duplicated(alltaxa$hname) | duplicated(alltaxa$hname,fromLast = TRUE),] %>% arrange(hname)
taxa.notdup <- anti_join(alltaxa,taxa.dup,by="hname")

# write csv
write.csv(taxa.dup,file="taxa.dup.out.csv")

taxa.dup.in <- read.csv("taxa.dup.in.csv")

# merge new and old host names
host.taxa <- bind_rows(taxa.notdup,taxa.dup.in) %>% arrange(hname)
host.taxa$H_F <- toupper(host.taxa$H_F)
host.taxa$H_O <- toupper(host.taxa$H_O)

# correct spelling errors
host.taxa$H_O[which(host.taxa$H_O == "TETRADONTIFORMES")] <- "TETRAODONTIFORMES"
host.taxa$H_F[which(host.taxa$H_F == "TETRADONTIDAE")] <- "TETRAODONTIDAE"
host.taxa$H_F[which(host.taxa$H_F == "CENTROLOPPHIDAE")] <- "CENTROLOPHIDAE"
host.taxa$H_F[which(host.taxa$H_F == "CLUPIDAE")] <- "CLUPEIDAE"
host.taxa$H_F[which(host.taxa$H_F == "OSTEOGOGLOSSIDAE")] <- "OSTEOGLOSSIDAE"
host.taxa$H_O[which(host.taxa$H_F == "SALMINUS")] <- "CHARACIFORMES"
host.taxa$H_F[which(host.taxa$H_F == "SALMINUS")] <- "BRYCONIDAE"
# just remove this one, and from host.traits & pairs above
host.taxa <- host.taxa[-which(host.taxa$hname=="Colossoma_macroponum"),] 

save(host.taxa,file="host.taxa.RData")

```

Generalism of the parasites can be calculated in a variety of ways, and may take into account relative prevalence, abundance and/or relatedness between host species [Poulin & Mouillot 2003, 2005]. Other indices account for connectedness within a host-parasite network and may account for prevalence or abundance [Dormann].  

In this data set, we don't have information on number of parasites per host or on prevalence or study effort, only reported associations between hosts and parasites. As a result we calculate only:

* **Degree**: the number of host species a parasite is associated with
* **$S_{TD}$**: the average taxonomic distinctness of all host species used by a parasite species [Poulin & Mouillot 2003]: $$S_{TD}=2[s(s-1)]^{-1}\sum \sum_{i<j} \omega_{ij}$$

    $S_{TD}$ can only be calculated for parasites infecting more than one host species, those infecting one species are assigned a value of 1. In addition $VarS_{TD}$ can only be calculated for hosts infecting more than 2 host species:
$$VarS_{TD}=[s(s-1)]^{-1}\sum \sum_{i\neq j}\left (\omega_{ij}-S_{TD}  \right )^2$$
    But this is the same as: $$VarS_{TD}=2[s(s-1)]^{-1}\sum \sum_{i< j}\left (\omega_{ij}-S_{TD}  \right )^2$$





* Another method could use host phylogenetic distance (eg from time tree of life) instead of categorized phylogenies
    + Do we have data for all species? No

```{r,cache=TRUE}
library(ape,suppressPackageStartupMessages())

tree <- read.tree("C:/Users/Josephine/Google Drive/Thesis/Chapter 3 (Community Ecology)/ch3 stuff/TimetreeOfLife2015.nwk")

H_bin <- paste(dat.u$H_Genus,dat.u$H_species,sep="_") %>% unique()

pruned.tree<-drop.tip(tree, setdiff(tree$tip.label, H_bin))

# calculate path distance between each species (this should be time since last common ancestor*2)
pathlength <- cophenetic.phylo(pruned.tree)

paste("number of host species in database:",length(H_bin),sep=" ")
paste("number of nodes in pruned tree:", length(pruned.tree$tip.label),sep=" ")


```


Divergence time data are only available for about a quarter of the host species in the fish database. 

Calculate bipartite indices


Calculate S_TD using taxonomic distance from data

```{r,eval=FALSE}
load("~/dat/generalismtheory/pairs.traits.RData")
load("~/dat/generalismtheory/host.taxa.RData")

# calculate S_TD from separate data frames

# # http://stackoverflow.com/questions/17171148/non-redundant-version-of-expand-grid
expand.grid.unique <- function(x, y, include.equals=FALSE)
{
    x <- unique(x)

    y <- unique(y)

    g <- function(i)
    {
        z <- setdiff(y, x[seq_len(i-include.equals)])

        if(length(z)) cbind(x[i], z, deparse.level=0)
    }

    do.call(rbind, lapply(seq_along(x), g))
}

# write function to calculate S_TD
# only congeneric hosts will have value 1, so call one species only value 0?

get.S_TD <- function(dat) {
  S_TD <- 0
  VarS_TD <- NA
  
  if(nrow(dat) > 1) {
    temp <- data.frame(expand.grid.unique(dat$hname,dat$hname))
    temp$genus <- 0
    temp$family <- 0
    temp$order <- 0
    temp$class <- 0
    temp$phylum <- 0
    
    for (i in 1:nrow(temp)) {
      temp2 <- rbind(dat[dat$hname == temp$X1[i],],dat[dat$hname == temp$X2[i],])
      temp$genus[i] <- as.numeric(identical(temp2$H_Genus[1],temp2$H_Genus[2]))
      temp$family[i] <- as.numeric(identical(temp2$H_F[1],temp2$H_F[2]))
      temp$order[i] <- as.numeric(identical(temp2$H_O[1],temp2$H_O[2]))
      temp$class[i] <- as.numeric(identical(temp2$H_class[1],temp2$H_class[2]))
      temp$phylum[i] <- as.numeric(identical(temp2$H_Phylum[1],temp2$H_Phylum[2]))
    }

    temp$omega <- with(temp, 6 - genus - family - order - class - phylum)

    S_TD <- 2*(sum(temp$omega))/(nrow(dat)*(nrow(dat)-1))
    temp$Var <- (temp$omega - S_TD)^2 
    #print(temp)
    VarS_TD <- 2*(sum(temp$Var))/(nrow(dat)*(nrow(dat)-1))
  }
  return(data.frame(S_TD=S_TD,VarS_TD=VarS_TD,degree=nrow(dat)))
  #return(paste(S_TD,VarS_TD,sep="@"))
}

pairswithtaxa <- pairs.traits %>% select(hname,pname) %>% left_join(host.taxa,by=("hname"))
# test <- pairswithtaxa[which(pairswithtaxa$pname=="Acanthocephaloides_distinctus"),]
# get.S_TD(test)

index <- pairswithtaxa %>% group_by(pname) %>% do(get.S_TD(.)) %>% ungroup()
save(index,file="index.RData")
load("para.traits.RData") # still has 4 duplicates



````

The triangular scatter of the index matches what was observed in Poulin & Mouillot 2003.

```{r,cache=TRUE}
load("index.RData")
plot(x=index$degree,y=index$S_TD)
plot(x=log10(index$degree),y=index$S_TD)

#boxplots by parasite characteristic
fulldat <- full_join(degree,index,by=c("P_bin","degree"))
# dim correct

boxplot(fulldat$S_TD ~ fulldat$Gender)
boxplot(fulldat$S_TD ~ fulldat$Host_stage)
boxplot(fulldat$S_TD ~ fulldat$Endoparasite)
boxplot(fulldat$S_TD ~ fulldat$Complex)
boxplot(fulldat$S_TD ~ fulldat$Sexual)
boxplot(fulldat$S_TD ~ fulldat$Reproduction)
boxplot(fulldat$S_TD ~ fulldat$Host_no)
boxplot(fulldat$S_TD ~ fulldat$Environment_LHS)
boxplot(fulldat$S_TD ~ fulldat$Env_mode)
boxplot(fulldat$S_TD ~ fulldat$Motile_LHS)
boxplot(fulldat$S_TD ~ fulldat$Vector)
boxplot(fulldat$S_TD ~ fulldat$Trophic)
boxplot(fulldat$S_TD ~ fulldat$Horizontal)
boxplot(fulldat$S_TD ~ fulldat$P_Taxon)


```

Does degree vary by taxonomic group or other parasite characteristics?

```{r,cache=TRUE}
boxplot(log(degree$degree) ~ degree$Gender)
boxplot(log(degree$degree) ~ degree$Host_stage)
boxplot(log(degree$degree) ~ degree$Endoparasite)
boxplot(log(degree$degree) ~ degree$Complex)
boxplot(log(degree$degree) ~ degree$Sexual)
boxplot(log(degree$degree) ~ degree$Reproduction)
boxplot(log(degree$degree) ~ degree$Host_no)
boxplot(log(degree$degree) ~ degree$Environment_LHS)
boxplot(log(degree$degree) ~ degree$Env_mode)
boxplot(log(degree$degree) ~ degree$Motile_LHS)
boxplot(log(degree$degree) ~ degree$Vector)
boxplot(log(degree$degree) ~ degree$Trophic)
boxplot(log(degree$degree) ~ degree$Horizontal)
boxplot(log(degree$degree) ~ degree$P_Taxon)
```

